---
title: "Homework 3"
author: "Robert Gambrel"
date: "May 31, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Q1

I've been trying to lose a few pounds this summer. A lot of guidance online suggests checking weight daily but looking only at weekly averages, since daily fluctuations can be drastic and not reflective of actual fat loss. How much water/food was ingested recently, how much salt, how recently you used the bathroom, and other factors give each weigh-in a lot of randomness around the 'true' value. I would therefore use a low alpha - this would emphasize the long-term trend instead of the day-to-day fluctuations.

# Q2

```{r}
pacman::p_load(dplyr, readr, purrr, outliers, ggplot2, lubridate, tidyr, stats, smooth, tibble, magrittr)


temps <- read_tsv('temps.txt')

temps_sorted <- temps %>%
	gather(year, high_temp, 2:21) %>%
	group_by(year) %>%
	mutate(rn = 1:n()) %>%
	ungroup() %>%	
	arrange(year, rn) %>%
	select(high_temp)

series <- ts(temps_sorted$high_temp, frequency = 123)

m <- HoltWinters(x = series, seasonal = 'mult', beta = F)
plot(m)
plot(fitted(m))

m <- es(series, model = "AAM")
plot(m)
plot(fitted(m))
m$model
m$formula
m$initial
m$initialSeason
fit <- as.tibble(m$fitted)
fit$year <- 0
```

```{r}
fit %<>%
  mutate(
    n = 1:n()
  )

for (i in 1:20) {
  fit %<>%
    mutate(
      year = ifelse((i - 1)*123 < n & n <= i*123, i, year)
    )
}

fit %<>%
  group_by(year) %>%
  mutate(
    rn = 1:n()
  )

ggplot(data = fit) +
  geom_line(aes(x = rn, y = `Series 1`, col = year, group = year))
```
# Q4
```{r}


crime <- read_tsv("http://www.statsci.org/data/general/uscrime.txt")


model <- lm(Crime ~ M + So + Ed + Po1 + Po2 + LF + M.F + Pop +
			  NW + U1 + U2 + Wealth + Ineq + Prob + Time, 
			  data = crime)
summary(model)

model2 <- lm(Crime ~ M + Ed + Po1 + Po2 +
			  U1 + U2 + Ineq + Prob, 
			  data = crime)
summary(model2)

AIC(model)
AIC(model2)

BIC(model)
BIC(model2)

new_data <- data.frame(M = 14.0, Ed = 10.0, Po1 = 12.0, Po2 = 15.5,
			     U1 = 0.120, U2 = 3.6, Ineq = 20.1, Prob = 0.04)

predict(model2, new_data)
```
